---
title: "Who Cares About Crackdowns? Exploring the Role of Trust in Individual Philanthropy"
short-title: "Who Cares About Crackdowns?"
author:
- name: Suparna Chaudhry
  affiliation: Lewis & Clark College
  url: http://www.suparnachaudhry.com/
  email: schaudhry@lclark.edu
- name: Marc Dotson
  affiliation: Brigham Young University
  url: https://marriottschool.byu.edu/directory/details?id=50683
  email: marc_dotson@byu.edu
- name: Andrew Heiss
  affiliation: Georgia State University
  url: https://www.andrewheiss.com/
  email: aheiss@gsu.edu
date: "October 2, 2020"
published: Working paper.
git-repo: https://github.com/andrewheiss/who-cares-about-crackdown
abstract: >-
  The phenomenon of closing civic space has adversely impacted the operations of INGOs. It has not just resulted in net losses of income to INGOs, but also how and where funds are allocated. We argue that individual private donors can be important in sustaining the operations of INGOs working in repressive contexts. Individual donors do not use the same performance-based metrics as official aid donors, and trust has shown to be an important component of support for nonprofits working towards difficult goals. How does trust in charitable organizations influence individuals’ preferences to donate, especially when these groups face crackdown? Using a simulated market for philanthropic donations towards INGOs, based on data from a nationally representative sample of individuals in the United States who regularly donate to charity, we find that trust in charitable organizations matters substantially in shaping donor preferences. Donor profiles with high levels of social trust are likely to donate to INGOs with friendly relationships with host governments. This support holds steady if INGOs face criticism or crackdown. In contrast, donor profiles with lower levels of social trust prefer to donate to organizations that do not face criticism or crackdown abroad. The global crackdown on NGOs may thus possibly sour NGOs' least trusting individual donors. Our findings have practical implications for INGOs raising funds from individuals amid closing civic space.
toc: false
reference-section-title: References
bibliography: bibliography.bib
title-page: true
word-count: "7,562"
keywords: international NGOs, donor behavior, philanthropy, human rights, repression, international giving, donor choice
thanks: >-
  The data that support the findings of this study, as well as the code to reproduce the findings, are openly available at <https://github.com/andrewheiss/who-cares-about-crackdown> and <https://osf.io/sm5ew/>. This experiment received approval by the human subjects research committees at Christopher Newport University (1436622-1), Brigham Young University (E19135), and Georgia State University (H19644). As part of a larger project, we preregistered hypotheses and our research design at the Open Science Framework, and our preregistration protocol is available at <https://osf.io/hsbyd/>. This study covers the third research question of our preregistered design.
author-note: >-
  Suparna Chaudhry, PhD, Assistant Professor of International Affairs, Department of International Affairs, Lewis \& Clark College; Marc Dotson, PhD, Assistant Professor of Marketing, Marriott School of Business, Brigham Young University; Andrew Heiss, PhD, Assistant Professor of Public Management, Andrew Young School of Policy Studies, Georgia State University. 
correspondence: >-
  Correspondence concerning this article should be addressed to Andrew Heiss, PhD, Department of Public Management and Policy, Andrew Young School of Policy Studies, Georgia State University, 14 Marietta Street NW, Suite 357, Atlanta, GA 30303. Contact: aheiss@gsu.edu
mainfont: Lora
sansfont: IBM Plex Sans
policy-implications: >-
  In the wake of the global crackdown on civil society, international NGOs (INGOs) need to maintain access to multiple sources of funding, including individual donors. 
  It is critical for INGOs to understand what motivates individuals to donate, especially if INGOs face legal crackdown in their host countries.
  
  Trust in political institutions and in charitable organizations matters substantially in shaping donor preferences.
  We find that potential donors with low social trust tend to be wary of negative host-country relationships and prefer to donate to INGOs that are friendly with their host governments, while potential donors with high social trust are more willing to stick with legally besieged INGOs.
  Donors that are the most trusting and involved may remain supportive when charities face government criticism and crackdown. The global crackdown on NGOs may also possibly sour NGOs' least trusting individual donors.
editor_options: 
  chunk_output_type: console
---

<!-- manuscript.md is generated from manuscript.Rmd. Only edit the .Rmd file, *not* the .md file. -->

```{r setup, include=FALSE}
# Figure out output format
is_docx <- knitr::pandoc_to("docx") | knitr::pandoc_to("odt")
is_latex <- knitr::pandoc_to("latex")
is_html <- knitr::pandoc_to("html")

# Word-specific things
table_format <- ifelse(is_docx, "huxtable", "kableExtra")  # Huxtable tables
conditional_dpi <- ifelse(is_docx, 300, 300)  # Higher DPI
conditional_align <- ifelse(is_docx, "default", "center")  # Word doesn't support align

# Knitr options
knitr::opts_chunk$set(
  echo = FALSE, warning = FALSE, message = FALSE,
  tidy.opts = list(width.cutoff = 120),  # Code width
  fig.retina = 3, dpi = conditional_dpi,
  fig.width = 7, fig.asp = 0.618,
  fig.align = conditional_align, out.width = "100%",
  fig.path = "output/figures/",
  cache.path = "output/_cache/",
  fig.process = function(x) {  # Remove "-1" from figure names
    x2 = sub('-\\d+([.][a-z]+)$', '\\1', x)
    if (file.rename(x, x2)) x2 else x
  }
)

# R options
options(
  width = 90,  # Output width
  dplyr.summarise.inform = FALSE,  # Turn off dplyr's summarize() auto messages
  knitr.kable.NA = "",  # Make NAs blank in kables
  kableExtra.latex.load_packages = FALSE,  # Don't add LaTeX preamble stuff
  modelsummary_default = table_format  # Set modelsummary backend
)
```

```{r libraries-data, include=FALSE}
library(tidyverse)
library(targets)
library(kableExtra)
library(huxtable)
library(patchwork)
library(scales)
library(janitor)
library(glue)
library(here)

withr::with_dir(here::here(), {
  source(tar_read(plot_funs))
  
  tar_load(personas)
  tar_load(orgs)
  tar_load(participant_summary)
  tar_load(sim_final)
  tar_load(sim_excel_final)
})
```

In 2016, Human Rights Watch claimed that civil society was under more aggressive attack than at any time in recent memory [@Roth:2016]. Governments across the globe have cracked down on civil society organizations (CSOs) and non-governmental organizations (NGOs)—a phenomenon known as "closing civic space" [@CarothersBrechenmacher:2014; @CIVICUS:2017; @DPR:2015; @DPR:2016].[^NGOdef] Legal restrictions, or what we refer to as legal crackdowns, are a core part of these efforts. These crackdowns create barriers to entry, funding, and advocacy for NGOs in an effort to control, obstruct, and repress these organizations. Barriers to funding are especially pervasive and restrict the ability of NGOs to secure financial resources. Restrictive states may prevent the transfer of foreign funds to NGOs based on the origin of these funds and which issues these funds can be used towards.

As a result of these laws, foreign aid (also referred to as public or official aid) channeled through NGOs has decreased considerably in repressive countries [@Brechenmacher:2017; @ChaudhryHeiss:2018; @DupuyPrakash:2018]. However, philanthropy from private donors and foundations is not as adversely affected [@McGill:2018]. Many foundations have continued channeling funds to countries with hostile legal environments [@NeedlesWolcheckSato:2018]. Given this trend of reduced public aid following restricted legal environments for NGOs abroad, INGOs may need to devote more resources to fundraise from individual donors. 

Individual private donations are a critical source of INGO funding. While total private philanthropy[^phildef] in the United States has grown substantially over time, it is notable that private giving to NGOs working in international affairs  and international affairs-focused organizations are one of only two categories of organizations that have continued experiencing substantial growth in giving and reached \$22.08 billion in 2018 [@givingusa:2019].[^intlaffairs] Most of this growth is driven by individuals making small-scale or micro-donations.[^smalldonors] In 2017, 70% of donations to international affairs NGOs came from individuals, rising to 86% in 2018 [@givingusa:2019]. But relying on individual donors means that NGOs must consider their preferences when fundraising. Which individual donor characteristics and attributes are most salient for INGOs interested in maximizing philanthropy at a time when they are facing increasingly hostile environments abroad? 

There exists a robust scholarship on the determinants of individual-level philanthropy to domestic causes. But the dynamics of *international* giving to NGOs are under-examined in the literature on individual donor attributes and preferences.[^exceptions] Motivating individuals to donate to international causes is hard as the number of recipients are large and far removed from the donor [@CasaleBaumann:2015, 100]. Trust is an essential component for INGO sustainability and for many nonprofits, trust often translates into grants and donations [@coombs2007]. Geography often impedes donors' direct evaluation of their work and trust is one of the most influential criteria that private donors use when deciding which nonprofits to support [@prakash2010trust; @bekkers2003trust]. A lack of trust could be further fueled by legal trouble with the NGO host government—a dynamic that individuals may not wish to enter with their dollars. How does trust in NGOs influence individuals' preferences to donate, especially when groups face crackdown? 

We explore this using a simulated market for philanthropic donations towards INGOs, based on data from a nationally representative sample of individuals in the United States who regularly donate to charity. We adapt methods used in marketing to generate dozens of distinct simulated personas that represent typical donor profiles, and we vary these personas along different dimensions of demographic, political, and social attributes. We find that an organization's target issue area and its relationship with its host government interact strongly with individuals' characteristics. Personas with high levels of trust in political and charitable institutions and who are actively engaged in charity and volunteerism prefer donating to human rights organizations that have friendly relationships with their host governments. In contrast, personas with low levels of social trust eschew both human rights organizations *and* any type of NGO with more contentious relationships with host governments. Instead, they prefer donating to NGOs with friendly host-country relationships and working on less contentious issues like emergency response and refugee relief. Personas with higher trust are more comfortable donating to legally besieged INGOs. Our findings imply that individual donors with high trust would continue remaining supportive of INGOs when they face criticism and crackdown from foreign governments. On the other hand, government crackdown may further sour NGOs least trusting individual donors.  

This study contributes both to scholarship and practice in public policy, nonprofits, and philanthropy in three important ways. It helps us understand how the international community and NGOs might respond to the phenomenon of closing civic space, which does not appear to be receding. Most NGOs working internationally have previously relied on government and foundation funding. But in the face of decreasing funds from official donors, our results show that INGOs can rely on individual donors, specifically, their most trusting donors, when groups face a hostile environment abroad. These findings lend support to previous literature arguing that trust matters  when individuals donate to organizations acting under great uncertainty or working towards difficult goals [@Wiepking:2010].  

Second, knowledge of demographic traits and its impact on funding can help nonprofits craft more effective and targeted marketing strategies [@Rajan:2009, 414]. Our paper shows that in readjusting their fundraising strategies towards individuals, framing, or the process through which actors present information to influence perceptions, is particularly important [@chong2007framing]. Framing affects their willingness to donate. However, existing research is unclear about which frames are effective when making appeals to individual donors considering international philanthropy. Since our study looks at multiple organizational and host country factors, our results can help NGOs frame their appeals. In particular, our results show that NGOs may benefit from publicizing crackdown at least towards their most trusting donors. 

Finally, while research in public policy examines giving to NGOs, philanthropy is not seen as relevant by international relations as a discipline [@youde2019]. However, philanthropy—particularly donations received by INGOs—can help us understand changes in the dynamics of global governance. Private individuals exercise power and influence in unique ways that differ from other non-state actors. As @youde2019 [p. 44] points out, it changes our conceptions of individuals being passive recipients to active entities that are not merely objects of governance. Thus, it can shed light on the changing logic of global governance.

Below, we first examine the state of the existing literature on the key drivers of individual giving. We then describe our survey experiment and simulation methods and explore the results. We conclude by discussing the implications of our findings for INGOs facing constrained space for civil society.


# Closing civic space and its impact on INGOs

Foreign aid to civil society—and NGOs in particular—has steadily increased over time. Many government and foundation donors perceive NGOs as nimbler and less bureaucratic than governments. In countries with weak institutions and poor governance, where direct aid transfers to governments raise issues of misuse and bureaucratic inefficiency, donors specifically seek out NGOs for projects [@Dietrich2013]. However, NGO-state relationships have evolved over time—in the 1950s and 60s, international assistance channeled through NGOs aimed at producing socio-economic change rather than political progress. But in the aftermath of the dissolution of the Soviet Union, many NGOs had the explicit goal of promoting democracy and bringing down authoritarian governments. Subsequently, many states perceive them as threatening and have sought to repress them.

Mre than 100 countries have proposed or enacted 244 measures restricting, repressing, or shutting down civil society since 2013 (ICNL 2021 CITATION FROM ISA PAPER). This has had a three-hold impact on INGOs. First, it has led to a reduction in resources and grants available to INGOs. Second, restrictive NGO laws can affect INGO programming by preventing organizations from working in certain (contentious) issue areas. To deal with this, some groups have adapted by changing the nature of their programming to appear less threatening to the repressive countries. Finally, in the worst-case scenario, this has led to a loss in INGOs' access to target countries—often also associated with the use of violence against INGOs and their employees. In this paper, we exlusively focus on the first dilemma—the reduction in INGO funding and possible responses to it. 


## Changes in INGO funding

Many INGOs in the Global South receive funds from a variety of public and private sources each year. Public funds—or traditional donor aid—channel money from official aid agencies through a variety of bilateral and multilateral institutions. Private sources include foundations, private corporations, and individuals. While INGOs have not been passive in responding to the crackdown on civil society, they face increasing obstacles from a funding perspective. 

Closing civic space not just resulted in net losses of income to INGOs, particularly through official bilateral and multilateral aid [@DupuyPrakash:2018], it has also affected how and where funds are allocated. Following the passage of anti-NGO laws, official donors move funds away from contentious causes such as human rights, media, corruption, advocacy, among others, towards tamer and more regime-compatible causes such as health, education [@ChaudhryHeiss:2018]. The latter trend has also been exacerbated by a preference among donors working in official aid and the democracy promotion establishment to gradually increase funding directed towards tame causes, at the expense of more contentious causes. This is because many tame causes can have easily quantifiable, measurable outputs, which makes it clear to show a program's effectiveness over time [@Bush:2015].

We argue that individual private donors can help fill in an important gap. Individual donors do not use the same performance-based metrics as official donors [@desaikharas]. Rather, individual donors often use heuristics to simply their decision-making, making cursory judgments about an organization's issue area, mission, vision, and values and seek out supplementary information from friends, family, and acquaintances [@TBPrakash:2017; @SzperPrakash:2011]. Due to these differences, individual donors can help address some of the challenges arising from reduced public funding available to INGOs. To be clear, we are not arguing that giving by individual donors would be a perfect substitute for NGOs net loss of income or income for certain programs, but rather, an important supplement to reduced funding from foreign sources. Moreover, it would also be a supplement to private funding from foundations towards international nonprofits which has not been as adversely affected, with foundations continue to channel funds to countries with hostile legal environments [@NeedlesWolcheckSato:2018].

However, research on the determinants of individual-level philanthropy for international causes—particularly towards organizations working in challenging environments—is still in its infancy due to the absence of expansive and high-quality data [@greenhill2013]. Moreover, a majority of research on giving to international groups is restricted to elite high net-worth donors[^highnetworth] [@USTrust2014]. We still know little about demographic traits and experiences that motivate or deter individuals to donate during shrinking civic space. Below, we look at key drivers of individual-level philanthropy and show how the importance of trust in donating during civic crackdowns. 

## Key drivers of individual-level philanthropy

Substantial research has looked at the individual-level determinants of donating to local, rather than international causes, including four main attributes: social capital, the role of higher education, levels of religiosity, and trust.

Social capital, particularly associational capital, or "the networks of formal and informal relationships to which people are associated," is often considered an important indicator of giving [@havens2007, 240]. Volunteering exposes individuals to the need for donations, making them feel more compelled to donate [@hossainlamb2012]. Those who participate in a variety of associations are both more likely to donate as well as make larger donations [@hossainlamb2017].

Higher education also influences charitable-giving decisions as it fosters prosocial motivations. Education brings people into social networks that entail a higher level of solicitation, which is vital in giving to international causes [@BekkersWiepking:2011]. Education also helps determine the kinds of causes that people donate to globally. @Micklewright:2009 show that higher education matters more for donating to international relief organizations than donating to domestic organizations. 
 
A number of NGOs facing crackdowns abroad also work on issues that may be seen as "contentious," including human rights, elections, corruption, advocacy, and media freedom—issues that may have a harder time attracting funds due to their non-essential nature when compared to causes like health, education, sanitation, and social services [@Bush:2015]. However, additional education has been shown to increase abstract thinking [@wiepkingmaas, 1978], which may be necessary to donate to causes that do not have quantifiable goals, including politically sensitive or contentious causes mentioned above.

Religiosity is also a significant predictor of the likelihood and level of giving. People who regularly attend religious services are more likely to have prosocial values, face a stronger norm of charitable giving, and may feel pressure to conform to group and community standards [@WiepkingBekkers:2010]. Regular attendance may also strengthen individuals' beliefs in organizations' missions.

Finally, an individual's disposition to trust, or the general tendency to trust others, including individuals groups or institutions has been shown to have an impact on both intention to donate as well as actual donation behavior. Trusting people are not only more likely to give to charitable causes, but also donate more [@BrownFerris2007]. 

Which of these characteristics matter when deciding to donate to an INGO, particularly when an organization faces legal crackdown abroad? Overall, individuals with higher levels of income, education, and greater religious proclivities are not only more likely to give internationally, but also give higher amounts [@Rajan:2009; @CasaleBaumann:2015]. Additionally, we posit that social trust can play an important role in the decision to donate, as individuals who trust NGOs are less likely than others to halt their monetary support in the face of difficulties [@dwyer1987]. 

We explore this question using a conjoint experiment conducted with a nationally representative sample in the U.S. Importantly, because there is so little research about international donations during legal crackdown by host governments, our analysis and results are exploratory and descriptive. Accordingly, we do not test specific hypotheses or theories. 

TODO: Fix final point - connect testing stuff with existing theories - we don't have a list of things to test

However, we do talk about causal effects given that we conduct an experiment that varies a number of factors in order to elicit opinions from respondents. Instead, we structure our analysis around broader themes from the existing literature and investigate how donors' (1) demographic characteristics, (2) public affairs knowledge and experience, and (3) levels of trust in government, charitable work, and associational life change individual donation preferences across different organizational issue areas, relationships with host governments, and funding sources. Our innovative experimental approach provides rich quantitative details of individual and organizational characteristics simultaneously and reveals unique insights into donor motivations that both confirm and expand previous research.


# Data and methods

## Conjoint experiment

We use a conjoint experiment to study the simultaneous effects of organizational characteristics and individual attributes on individuals' preferences to donate to a nonprofit organization. Conjoint analysis allows us to elicit donor preferences without respondents explicitly stating their exact preferences. This kind of analysis is commonly used in marketing to reveal consumer preferences for hypothetical products, and has become increasingly popular in political science and public policy [@Helfer:2016; @Mummolo:2016; @KnudsenJohannesson:2019]. Because we are interested in exploring the complexity behind the decision to donate, we apply conjoint methods to the study of nonprofits and philanthropy.

In our experiment, we presented respondents with four randomly selected international nonprofit organizations with four randomly assigned features.[^appendix-survey] We included organizations that are associated with a range of contentious and noncontentious issues: Amnesty International, Greenpeace, Oxfam, and the International Committee of the Red Cross. Because of their strong name recognition, these organizations (and others like World Vision, the World Wildlife Fund, and Médecins Sans Frontières) are commonly used in experimental surveys on philanthropy [@FaulknerTruongRomaniuk:2015].

\stgroup

\singlespacing

```{r organization-attributes-full}
tbl_org_attributes_caption <- "Organization attributes varied in the experiment. Each participant saw repeated combinations of attributes, with one attribute randomly selected from each column."

tbl_org_attributes <- orgs$org_attributes %>% 
  select(Organization, `Issue area`, `Organizational practices`, 
         `Funding sources`, `Relationship with government`)

if (!is_docx) {
  tbl_org_attributes %>%
    kbl(align = c("l", "l", "l", "l", "l"),
        booktabs = TRUE, linesep = "\\addlinespace",
        caption = tbl_org_attributes_caption) %>% 
    kable_styling() %>% 
    column_spec(1:5, width = "0.17\\\\textwidth")
} else {
  tbl_org_attributes %>% 
    hux() %>% 
    set_font_size(10) %>%
    set_all_padding(3) %>% 
    set_top_border(1, everywhere) %>%
    set_bottom_border(c(1, 5), everywhere) %>%
    set_width(1) %>% 
    set_wrap(TRUE) %>% 
    huxtable::set_caption(tbl_org_attributes_caption) %>% 
    huxtable::as_flextable() %>% 
    flextable::font(fontname = "Times New Roman")
}
```

\fingroup

The hypothetical organizations are based on random combinations of features, including one of four issue areas, the presence or absence of two organizational practices, one of three funding sources, and one of three relationships with host governments (see Table \@ref(tab:organization-attributes-full); see also Q3.1 in the survey text in the appendix). Respondents saw twelve sets of four hypothetical organizations out of 288 (4 × 4 × 2 × 3 × 3) possible combinations, meaning no single respondent was offered every combination. While not every combination necessarily reflected reality—for instance, a respondent might have been asked to compare a fictional Amnesty International that focuses on environmental issues and that is funded by private donations with a Greenpeace that focuses on human rights and that is funded by government grants—it is not important that these features align with true organizational characteristics, since the goal of these repeated hypothetical questions is to identify which attributes are the most salient for donors. 

In June 2019, we recruited survey respondents through Centiment, a commercial provider of high-quality, non-probability opt-in survey panels. Centiment ensures panel quality by actively recruiting and paying representative samples of the U.S. population. To participate in the study, respondents were first screened based on their charitable activities. Those who indicated that they gave to charity once every few years or never were disqualified. After screening, we collected a sample of 1,016 respondents, which is sufficient for model estimation.[^sample-size]

\stgroup

\singlespacing

```{r sample-details}
tbl_participant_summary_caption <- "Summary of individual respondent characteristics"

tbl_participant_summary <- participant_summary %>% 
  select(Question = clean_name,
         Response = level,
         N = count,
         `%` = nice_prop) 

if (!is_docx) {
  tbl_participant_summary %>% 
    kbl(align = c("l", "l", "c", "c"),
        longtable = TRUE, booktabs = TRUE, linesep = "",
        caption = tbl_participant_summary_caption) %>% 
    pack_rows(index = table(fct_inorder(participant_summary$category)),
              latex_gap_space = "1.5em", indent = FALSE) %>%
    collapse_rows(columns = 1, valign = "top", latex_hline = "major") %>% 
    kable_styling(latex_options = c("repeat_header"))
} else {
  tbl_participant_summary %>% 
    hux() %>% 
    set_font_size(10) %>%
    set_all_padding(3) %>% 
    set_align(everywhere, 1:2, "left") %>% 
    set_align(everywhere, 3:4, "center") %>% 
    merge_repeated_rows(col = 1) %>% 
    insert_row("Demographics", "", "", "", after = 1) %>% 
    insert_row("\nAttitudes toward charity", "", "", "", after = 23) %>% 
    insert_row("\nPolitics, ideology, and religion", "", "", "", after = 51) %>%
    merge_cells(2, 1:2) %>%
    merge_cells(24, 1:2) %>%
    merge_cells(52, 1:2) %>%
    set_bold(1, everywhere) %>%
    set_bold(c(2, 24, 52), everywhere) %>%
    set_top_border(1, everywhere) %>%
    set_bottom_border(1, everywhere) %>%
    set_bottom_border(c(7, 9, 14, 21, 23,
                        26, 33, 40, 47, 51,
                        55, 57, 59, 66, 73, 75, 78, 80), 
                      everywhere) %>% 
    set_width(1) %>% 
    set_col_width(1, 0.25) %>%
    set_col_width(2, 0.4) %>%
    set_col_width(3:4, 0.1) %>%
    set_wrap(TRUE) %>% 
    huxtable::set_caption(tbl_participant_summary_caption) %>% 
    huxtable::as_flextable() %>% 
    flextable::font(fontname = "Times New Roman")
}
```

\fingroup

Table \@ref(tab:sample-details) provides a summary of the individual characteristics of survey respondents. Because we limited the sample to people who regularly make charitable contributions, most respondents have a favorable opinion of nonprofits and charities, with more than half reporting a 5 or greater level of trust in charities (on a 7-point scale). Moreover, nearly 90% indicated that it is important to trust charities. Additionally, the majority of respondents volunteer and donate to charity at least once a year and give between \$1–\$999 annually. Most rarely attend church, but feel that religion is important. Respondents are also civically engaged (three quarters voted in the last election) and follow national news closely (70% follow the news daily). Respondents are fairly balanced ideologically, with the majority reporting a 3, 4, or 5 on a 7-point scale of liberal–conservative political ideology.

We analyze the effect of different combinations of organizational features and individual respondent attributes with a two-level hierarchical Bayesian multinomial logit model.[^model-details] Simply put, the first level of the model predicts individual donor preferences for various combinations of features (e.g. the combination of human rights issues, financial transparency, government funding, and government crackdown), while the second level predicts the heterogeneity of these fitted preferences as a function of individual donor characteristics.

## Simulated philanthropy market

In this paper, we look at which kinds of donors are more or less likely to consider an NGO's issue area, funding sources, and relationships with host governments when deciding to donate. Disentangling the effect of every individual level-characteristic measured in the survey on every combination of organizational characteristics presented in the conjoint experiment is unwieldy and near impossible. Given the organizational attributes presented in our experiment, and the possible individual characteristics that respondents could select, there are billions of possible combinations and iterations of organizational and individual attributes—many of which rarely occur in the actual population of potential donors. To more fully explore the relationships between different organizational- and individual-level attributes, we turn to simulation methods that allow us to examine the impact of multiple covariates simultaneously. To simplify the possible combinations of individual characteristics, we generate 64 distinct personas with attributes that vary along different dimensions of demographics, politics and public affairs experience, and social attitudes (see Table \@ref(tab:persona-attributes)). We also generate 24 simulated organizations that vary by issue area, relationship with host government, and source of funding (see Table \@ref(tab:organization-attributes)).

\stgroup

\singlespacing

```{r persona-attributes}
tbl_persona_attributes_caption <- "Individual attributes varied in the simulation, resulting in 64 persona profiles"

tbl_persona_attributes <- personas$persona_attributes 

if (!is_docx) {
  tbl_persona_attributes %>% 
    kbl(align = c("l", "l", "l"),
        booktabs = TRUE, linesep = "\\addlinespace",
        caption = tbl_persona_attributes_caption) %>% 
    kable_styling() %>% 
    column_spec(1, width = "0.3\\\\textwidth") %>% 
    column_spec(2, width = "0.3\\\\textwidth") %>% 
    column_spec(3, width = "0.3\\\\textwidth")
} else {
  tbl_persona_attributes %>% 
    hux() %>% 
    set_font_size(10) %>%
    set_all_padding(3) %>% 
    set_align(everywhere, everywhere, "left") %>% 
    set_top_border(1, everywhere) %>%
    set_bottom_border(c(1, 9), everywhere) %>%
    set_width(1) %>% 
    set_wrap(TRUE) %>% 
    huxtable::set_caption(tbl_persona_attributes_caption) %>% 
    huxtable::as_flextable() %>% 
    flextable::font(fontname = "Times New Roman")
}
```

```{r organization-attributes}
tbl_org_attributes_caption <- "Organization attributes varied in the simulation, resulting in 24 hypothetical organizations"

tbl_org_attributes <- orgs$org_attributes %>% 
  select(`Issue area`, `Relationship with government`, `Funding`) 

if (!is_docx) {
  tbl_org_attributes %>% 
    kbl(align = c("l", "l", "l"),
        booktabs = TRUE, linesep = "\\addlinespace",
        caption = tbl_org_attributes_caption) %>% 
    kable_styling()
} else {
  tbl_persona_attributes %>% 
    hux() %>% 
    set_font_size(10) %>%
    set_all_padding(3) %>% 
    set_align(everywhere, everywhere, "left") %>% 
    set_top_border(1, everywhere) %>%
    set_bottom_border(c(1, 9), everywhere) %>%
    set_width(1) %>% 
    set_wrap(TRUE) %>% 
    huxtable::set_caption(tbl_org_attributes_caption) %>% 
    huxtable::as_flextable() %>% 
    flextable::font(fontname = "Times New Roman")
}
```

\fingroup

We then combine these different scenarios of donor and organization attributes with the coefficients from a multinomial multilevel regression model to calculate the probability that each simulated persona has of giving to specific combinations of organizational attributes. Any individual characteristics that we do not explicitly vary in Table \@ref(tab:persona-attributes) are held at their means or modal values from the survey sample. Table \@ref(tab:sim-output) provides an example of the simulation output, showing how two different personas are predicted to allocate donations to organizations with various characteristics. The proportion of donations from each persona sums to 100%, as every persona is guaranteed to make a donation.

\stgroup

\singlespacing

```{r sim-output}
example_personas <- c("persona2", "persona63")

example_persona_results <- sim_final %>% 
  filter(persona_id %in% example_personas) %>% 
  mutate(org_funding = str_to_sentence(str_remove(org_funding, "Mostly funded by "))) %>% 
  mutate(org_clean = glue("{organization}: {org_issue}, {org_funding}, {org_relationship}")) %>% 
  mutate(persona_desc = recode(
    persona_id,
    "persona2" = "Lower income high school graduate who rarely attends religious services; liberal who reads and travels; doesn't trust or donate",
    "persona63" = "Higher income college graduate who attends religious services; conservative who doesn't read or travel; trusts and donates")
  ) %>%
  mutate(persona_clean = glue("{persona}: {persona_desc}")) %>% 
  select(persona_clean, share, org_clean) %>% 
  pivot_wider(names_from = "persona_clean", values_from = "share") %>% 
  adorn_totals(where = "row", name = "Total")

example_persona_results_small <- example_persona_results %>% 
  slice(c(1, 2, 3, 7, 8, 9, 16, 17, 25)) %>% 
  mutate(across(where(is.numeric), ~ percent_format(accuracy = 0.1)(.))) %>% 
  add_row(org_clean = "…", .after = 3) %>% 
  add_row(org_clean = "…", .after = 7) %>% 
  add_row(org_clean = "…", .after = 10) %>% 
  mutate(across(everything(), ~replace_na(., "…"))) %>% 
  rename(Organization = org_clean)

example_personal_results_caption <- "Sample simulation output"

if (!is_docx) {
  example_persona_results_small %>% 
    # mutate(Organization = text_spec(Organization, bold = Organization == "Total")) %>% 
    kbl(align = c("l", "c", "c"),
        booktabs = TRUE, linesep = "\\addlinespace",
        caption = example_personal_results_caption) %>% 
    kable_styling() %>% 
    column_spec(1, width = "0.2\\\\textwidth") %>% 
    column_spec(2, width = "0.35\\\\textwidth") %>% 
    column_spec(3, width = "0.35\\\\textwidth")
} else {
  example_persona_results_small %>% 
    hux() %>% 
    set_font_size(10) %>%
    set_all_padding(3) %>% 
    set_align(everywhere, 1, "left") %>% 
    set_align(everywhere, 2:3, "center") %>% 
    set_top_border(1, everywhere) %>%
    set_bottom_border(c(1, 12, 13), everywhere) %>%
    set_width(1) %>% 
    set_wrap(TRUE) %>% 
    huxtable::set_caption(example_personal_results_caption) %>% 
    huxtable::as_flextable() %>% 
    flextable::font(fontname = "Times New Roman")
}
```

\fingroup

Relying on simulation methods offers useful analytic benefits. Rather than interpret individual marginal effects from a regression model, we can explore multiple moving parts from the same model simultaneously to understand how individual characteristics interact with each other. This allows us to more fully explore our key research questions and see which kinds of donors are more or less likely to consider specific organizational attributes when deciding to donate, and which donor characteristics are associated with the propensity to donate. Importantly, each persona is an artificial construct. While some survey respondents might match an exact persona profile, others may not. This is acceptable in simulations as each combination of persona characteristics is plausible in the real world.


# Results

These simulated findings are more abstract than the results from an ordinary survey sample. But by using simulations, we can discover interesting patterns of donor preferences when faced with organizations focusing on a variety of issues, funding sources, and legal restrictions abroad. Our market simulation reveals which combination of organizational features each persona is most likely to express a preference for donating to. We examine the results of the simulation in three stages, looking at how personas with different (1) demographic characteristics, (2) public affairs knowledge and experience, and (3) levels of trust in government, charitable work, and associational life change their donation preferences across different organizational issue areas, relationships with host governments, and funding sources.

## Patterns across demographics

We begin by looking at differences in donation preferences across a range of basic demographic characteristics. The four panels of Figure \@ref(fig:income) show the average predicted market share for different organizational issue areas, funding sources, and relationships with host governments across different levels of persona income, with two different cutpoints for high and low income. To ease with modeling and interpretation, we split high- and low-income personas based on the 2017 U.S. median income (\$61,372), showing personas with incomes higher and lower than the national median. Because there is wide variation in donor incomes in the U.S., we also show the results from a simulation model with a more extreme split in income, holding annual persona income at \$50,000 and \$100,000+. The difference in the division of high and low income makes no difference in the market shares across issue areas and funding sources, but there are substantial income-based differences across different relationships with host governments.

```{r income, fig.width=6, fig.height=4.5, fig.asp=NULL, out.width="100%", fig.cap="Average predicted donation market shares across all personas, segmented by persona income across organizational issue area, relationship with host government, and funding sources"}
plot_income_issue <- sim_final %>% 
  group_by(org_issue, persona_income) %>% 
  summarize(avg_share = mean(share)) %>% 
  mutate(facet = "Issue area") %>% 
  ggplot(aes(y = fct_rev(str_wrap_factor(org_issue, 15)),
             x = avg_share, color = fct_rev(persona_income))) +
  geom_pointrange(size = 0.75, fatten = 1.5,
                  aes(xmin = 0, xmax = ..x..), position = position_dodge(width = 0.5)) + 
  scale_x_continuous(labels = percent_format(accuracy = 1), expand = expansion(add = c(0, 0.002)),
                     breaks = seq(0, 0.06, 0.02)) +
  scale_color_manual(values = clrs_ngo_pairs[[1]], guide = FALSE) +
  coord_cartesian(xlim = c(0, 0.06)) +
  labs(x = "Average donation share", y = NULL, color = NULL) +
  facet_wrap(vars(facet)) +
  theme_ngo() +
  theme(panel.grid.major.y = element_blank())

plot_income_funding <- sim_final %>% 
  group_by(org_funding, persona_income) %>% 
  summarize(avg_share = mean(share)) %>% 
  mutate(facet = "Funding sources") %>% 
  ggplot(aes(y = fct_rev(str_wrap_factor(org_funding, 10)), 
             x = avg_share, color = fct_rev(persona_income))) +
  geom_pointrange(size = 0.75, fatten = 1.5,
                  aes(xmin = 0, xmax = ..x..), position = position_dodge(width = 0.5)) + 
  scale_x_continuous(labels = percent_format(accuracy = 1), expand = expansion(add = c(0, 0.002))) +
  coord_cartesian(xlim = c(0, 0.06)) +
  scale_color_manual(values = clrs_ngo_pairs[[1]], guide = FALSE) +
  labs(x = "Average donation share", y = NULL, color = NULL) +
  facet_wrap(vars(facet)) +
  theme_ngo() +
  theme(panel.grid.major.y = element_blank())

plot_income_relationship <- sim_final %>% 
  mutate(persona_income = fct_recode(persona_income,
                                     "< $61,372/year" = "Lower income",
                                     "> $61,372/year" = "Higher income")) %>% 
  group_by(org_relationship, persona_income) %>% 
  summarize(avg_share = mean(share)) %>% 
  mutate(facet = "Relationship with host government") %>% 
  ggplot(aes(y = fct_rev(str_wrap_factor(org_relationship, 10)), 
             x = avg_share, color = fct_rev(persona_income))) +
  geom_pointrange(size = 0.75, fatten = 1.5, 
                  aes(xmin = 0, xmax = ..x..), position = position_dodge(width = 0.5)) +
  scale_x_continuous(labels = percent_format(accuracy = 1), expand = expansion(add = c(0, 0.003))) +
  coord_cartesian(xlim = c(0, 0.1)) +
  scale_color_manual(values = clrs_ngo_pairs[[1]], 
                     guide = guide_legend(reverse = TRUE, nrow = 1,
                                          override.aes = list(size = 0.25,
                                                              linetype = 0))) +
  labs(x = "Average donation share", y = NULL, color = NULL) +
  facet_wrap(vars(str_wrap(facet, 50))) +
  theme_ngo() +
  theme(panel.grid.major.y = element_blank(),
        legend.key.width = unit(0.5, "lines"))

plot_income_relationship_extreme <- sim_excel_final %>% 
  mutate(persona_income = fct_recode(persona_income,
                                     "$50,000/year" = "Lower income",
                                     "$100,000/year" = "Higher income"),
         org_relationship = fct_recode(org_relationship,
                                       "Under crackdown" = "Crackdown")) %>% 
  group_by(org_relationship, persona_income) %>% 
  summarize(avg_share = mean(share)) %>% 
  mutate(facet = "Relationship with host government") %>% 
  ggplot(aes(y = fct_rev(str_wrap_factor(org_relationship, 10)), 
             x = avg_share, color = fct_rev(persona_income))) +
  geom_pointrange(size = 0.75, fatten = 1.5, linetype = "21", pch = 2,
                  aes(xmin = 0, xmax = ..x..), position = position_dodge(width = 0.5)) + 
  scale_x_continuous(labels = percent_format(accuracy = 1), expand = expansion(add = c(0, 0.002))) +
  coord_cartesian(xlim = c(0, 0.1)) +
  scale_color_manual(values = clrs_ngo_pairs[[1]], 
                     guide = guide_legend(reverse = TRUE, nrow = 1,
                                          override.aes = list(size = 0.25, 
                                                              linetype = 0))) +
  labs(x = "Average donation share", y = NULL, color = NULL) +
  facet_wrap(vars(str_wrap(facet, 50))) +
  theme_ngo() +
  theme(panel.grid.major.y = element_blank(),
        legend.key.width = unit(0.5, "lines"))

plot_income <- ((plot_income_issue + labs(x = NULL)) + 
  (plot_income_funding + labs(x = NULL))) / 
  (plot_income_relationship + plot_income_relationship_extreme)

plot_income
```

Regardless of income, personas respond better to human rights and refugee assistance organizations, followed closely by emergency response organizations. Environmental NGOs have the lowest market share across both levels of incomes. There are negligible differences across income levels—wealthier personas show a slight preference towards human rights and refugee organizations. This is possibly because income correlates with interest in public affairs and political activity—individuals who are wealthier arguably have more time and resources to follow the news and be continuously engaged in their communities and in other activist causes. There are no sizable differences in donation preferences by income across different forms of NGO funding—both levels of income tend to prefer organizations funded by government grants over small donors.

An organization's relationship with its host government does influence market shares across levels of income, but the size of this difference depends on the magnitude of income. In general, personas prefer to donate to organizations with friendly relationships with their host governments and avoid those under crackdown. Wealthier donor personas are more likely to prefer donating to organizations that are criticized by their host governments. This falls in line with their preference for human rights organizations, which have more contentious missions and are more likely to be targeted. This support, however, evaporates as crackdown intensifies. High-income personas are the least likely demographic to prefer donating to NGOs under crackdown. Their lower income counterparts, on the other hand, are generally unfazed by NGOs' relationships with government, maintaining the same average market share across all three types of host country relationships. Patterns in preferences for issue area and funding sources remain constant across both methods of categorizing income—crackdown only has an effect on very high-income personas. Income thus appears to influence preferences particularly for NGOs facing crackdown abroad—wealthy donors are perhaps fair-weather friends that are more likely to pull their support when an organization's work becomes more difficult.

```{r edu-religion, fig.width=6, fig.height=4.5, fig.asp=NULL, out.width="100%", fig.cap="Average predicted donation market shares across all personas, segmented by persona education and religiosity across organizational issue area, relationship with host government, and funding sources"}
# Education plots
plot_education_issue <- sim_final %>% 
  group_by(org_issue, persona_education) %>% 
  summarize(avg_share = mean(share)) %>% 
  mutate(facet = "Issue area") %>% 
  ggplot(aes(y = fct_rev(str_wrap_factor(org_issue, 15)), 
             x = avg_share, color = fct_rev(persona_education))) +
  geom_pointrange(size = 0.75, fatten = 1.5,
                  aes(xmin = 0, xmax = ..x..), position = position_dodge(width = 0.5)) + 
  scale_x_continuous(labels = percent_format(accuracy = 1), expand = expansion(add = c(0, 0.002))) +
  coord_cartesian(xlim = c(0, 0.06)) +
  scale_color_manual(values = clrs_ngo_pairs[[2]], guide = FALSE) +
  labs(x = "Average donation share", y = NULL, color = NULL) +
  facet_wrap(vars(facet)) +
  theme_ngo() +
  theme(panel.grid.major.y = element_blank())

plot_education_funding <- sim_final %>% 
  group_by(org_funding, persona_education) %>% 
  summarize(avg_share = mean(share)) %>% 
  mutate(facet = "Funding sources") %>% 
  ggplot(aes(y = fct_rev(str_wrap_factor(org_funding, 10)), 
             x = avg_share, color = fct_rev(persona_education))) +
  geom_pointrange(size = 0.75, fatten = 1.5,
                  aes(xmin = 0, xmax = ..x..), position = position_dodge(width = 0.5)) + 
  scale_x_continuous(labels = percent_format(accuracy = 1), expand = expansion(add = c(0, 0.002))) +
  coord_cartesian(xlim = c(0, 0.06)) +
  scale_color_manual(values = clrs_ngo_pairs[[2]], guide = FALSE) +
  labs(x = "Average donation share", y = NULL, color = NULL) +
  facet_wrap(vars(facet)) +
  theme_ngo() +
  theme(panel.grid.major.y = element_blank())

plot_education_relationship <- sim_final %>% 
  group_by(org_relationship, persona_education) %>% 
  summarize(avg_share = mean(share)) %>% 
  mutate(facet = "Relationship with host government") %>% 
  ggplot(aes(y = fct_rev(str_wrap_factor(org_relationship, 10)),
             x = avg_share, color = fct_rev(persona_education))) +
  geom_pointrange(size = 0.75, fatten = 1.5,
                  aes(xmin = 0, xmax = ..x..), position = position_dodge(width = 0.5)) + 
  scale_x_continuous(labels = percent_format(accuracy = 1), expand = expansion(add = c(0, 0.002))) +
  coord_cartesian(xlim = c(0, 0.1)) +
  scale_color_manual(values = clrs_ngo_pairs[[2]], 
                     guide = guide_legend(reverse = TRUE, ncol = 1,
                                          override.aes = list(size = 0.25,
                                                              linetype = 0)),
                     labels = label_wrap(15)) +
  labs(x = "Average donation share", y = NULL, color = NULL) +
  facet_wrap(vars(str_wrap(facet, 20))) +
  theme_ngo() +
  theme(panel.grid.major.y = element_blank(),
        legend.key.width = unit(0.5, "lines"))

# Religion plots
plot_religion_issue <- sim_final %>% 
  group_by(org_issue, persona_religion) %>% 
  summarize(avg_share = mean(share)) %>% 
  mutate(facet = "Issue area") %>% 
  ggplot(aes(y = fct_rev(str_wrap_factor(org_issue, 15)), 
             x = avg_share, color = fct_rev(persona_religion))) +
  geom_pointrange(size = 0.75, fatten = 1.5, pch = 17,
                  aes(xmin = 0, xmax = ..x..), position = position_dodge(width = 0.5)) + 
  scale_x_continuous(labels = percent_format(accuracy = 1), expand = expansion(add = c(0, 0.002))) +
  coord_cartesian(xlim = c(0, 0.06)) +
  scale_color_manual(values = clrs_ngo_pairs[[4]], guide = FALSE) +
  labs(x = "Average donation share", y = NULL, color = NULL) +
  facet_wrap(vars(facet)) +
  theme_ngo() +
  theme(panel.grid.major.y = element_blank())

plot_religion_funding <- sim_final %>% 
  group_by(org_funding, persona_religion) %>% 
  summarize(avg_share = mean(share)) %>% 
  mutate(facet = "Funding sources") %>% 
  ggplot(aes(y = fct_rev(str_wrap_factor(org_funding, 10)), 
             x = avg_share, color = fct_rev(persona_religion))) +
  geom_pointrange(size = 0.75, fatten = 1.5, pch = 17,
                  aes(xmin = 0, xmax = ..x..), position = position_dodge(width = 0.5)) + 
  scale_x_continuous(labels = percent_format(accuracy = 1), expand = expansion(add = c(0, 0.002))) +
  coord_cartesian(xlim = c(0, 0.06)) +
  scale_color_manual(values = clrs_ngo_pairs[[4]], guide = FALSE) +
  labs(x = "Average donation share", y = NULL, color = NULL) +
  facet_wrap(vars(facet)) +
  theme_ngo() +
  theme(panel.grid.major.y = element_blank())

plot_religion_relationship <- sim_final %>% 
  group_by(org_relationship, persona_religion) %>% 
  summarize(avg_share = mean(share)) %>% 
  mutate(facet = "Relationship with host government") %>% 
  ggplot(aes(y = fct_rev(str_wrap_factor(org_relationship, 10)), 
             x = avg_share, color = fct_rev(persona_religion))) +
  geom_pointrange(size = 0.75, fatten = 1.5, pch = 17,
                  aes(xmin = 0, xmax = ..x..), position = position_dodge(width = 0.5)) + 
  scale_x_continuous(labels = percent_format(accuracy = 1), expand = expansion(add = c(0, 0.002))) +
  coord_cartesian(xlim = c(0, 0.1)) +
  scale_color_manual(values = clrs_ngo_pairs[[4]], 
                     guide = guide_legend(reverse = TRUE, ncol = 1,
                                          override.aes = list(size = 0.25,
                                                              linetype = 0)),
                     labels = label_wrap(20)) +
  labs(x = "Average donation share", y = NULL, color = NULL) +
  facet_wrap(vars(str_wrap(facet, 20))) +
  theme_ngo() +
  theme(panel.grid.major.y = element_blank(),
        legend.key.width = unit(0.5, "lines"))

plot_education_religion <- (plot_education_issue + labs(x = NULL)) + 
  (plot_education_funding + labs(x = NULL)) + 
  (plot_education_relationship + labs(x = NULL)) + 
  guide_area() +
  plot_religion_issue + 
  plot_religion_funding + 
  plot_religion_relationship +
  plot_layout(guides = "collect", ncol = 4)

plot_education_religion
```

While education and religiosity are both important drivers of individual philanthropy, neither are associated with any sizable differences in persona donation patterns (as seen in Figure \@ref(fig:edu-religion)). Human rights issues attract a slightly larger market share of personas who are college graduates or who attend religious services at least monthly, while refugee issues are more likely to attract less religious personas, but these differences are minor. Preferences are consistent for different sources of funding as well—personas generally prefer organizations funded mostly by government grants. Persona preferences across different organizational relationships with host governments are also consistent across education and religiosity, with most personas preferring to donate to NGOs that enjoy friendly relations abroad. 

## Patterns across public affairs knowledge, experience, and social trust

We next look at how differences in public affairs knowledge, experience, and political ideology influence a persona's propensity to donate to an organization. We vary personas along two dimensions: (1) political ideology, with semi-liberal and semi-conservative political leanings (2 and 5 on a 7-point scale), and (2) knowledge and experience with public affairs and the news, with personas that either follow national and international news and that have travelled abroad or personas that do not follow the news or travel. 

We combine these dimensions with an additional set of donor characteristics that vary a persona's level of social trust. Personas with high levels of trust have high trust in political institutions and charities, think people should be more charitable, frequently volunteer, donate to charity once a month, have a history of political activism, and are members of an association. Personas with low levels of trust have the opposite characteristics: they do not trust political institutions or charities, rarely donate or volunteer, and are not members of associations. 

```{r issue-social, fig.width=6, fig.height=4, fig.asp=NULL, out.width="100%", fig.cap="Average predicted donation market shares across all personas, segmented by persona public affairs knowledge, political ideology, and social trust across different NGO issue areas"}
sim_final %>% 
  group_by(org_issue, persona_trust, persona_ideology, persona_experience) %>% 
  summarize(avg_share = mean(share)) %>% 
  ggplot(aes(y = fct_rev(str_wrap_factor(org_issue, 15)), 
             x = avg_share, color = persona_trust)) +
  geom_pointrange(size = 0.75, fatten = 1.5, 
                  aes(xmin = 0, xmax = ..x..), position = position_dodge(width = 0.5)) + 
  scale_x_continuous(labels = percent_format(accuracy = 1), expand = expansion(add = c(0, 0.002))) +
  scale_color_manual(values = clrs_ngo_pairs[[3]], 
                     guide = guide_legend(reverse = TRUE,
                                          override.aes = list(size = 0.25,
                                                              linetype = 0))) +
  labs(x = "Average share of donations", y = NULL, color = NULL) +
  facet_grid(rows = vars(persona_ideology), cols = vars(str_wrap(persona_experience, 100))) +
  theme_ngo() +
  theme(panel.grid.major.y = element_blank(),
        legend.key.width = unit(0.5, "lines"))
```

Figures \@ref(fig:issue-social) and \@ref(fig:relationship-social) show the results of the market simulation across these three dimensions of persona characteristics, varying public affairs knowledge, political ideology, and social trust. Personas respond differently to organizational issue areas. In Figure \@ref(fig:issue-social), those who follow the news and travel abroad are more likely to prefer donating to human rights organizations than their less knowledgeable and less traveled counterparts. Personas who do not follow the news or travel show a greater preference for emergency response and refugee relief organizations. There are no strong ideological effects across issue areas. Both conservative and liberal personas tend to follow similar patterns of preference across NGO issue. 

Social trust plays a substantial role in determining simulated market share, however. Those who trust social and charitable institutions are very likely to prefer donating to human rights organizations, regardless of their public affairs knowledge or political leanings. In all circumstances, those with high levels of trust have between a 7–10% chance of donating to a human rights organization, on average. Personas with low levels of trust, on the other hand, strongly avoid human rights organizations and prefer instead to donate to refugee and emergency response NGOs. The difference is substantial—a liberal persona with high levels of public affairs knowledge and experience will give to a human rights organization 10% of the time, on average if they have high levels of trust, but less than 3% of the time if they are less trusting. Trust in charities and institutions is a key factor in deciding to donate to more contentious human rights organizations.

```{r relationship-social, fig.width=6, fig.height=4, fig.asp=NULL, out.width="100%", fig.cap="Average predicted donation market shares across all personas, segmented by persona public affairs knowledge, political ideology, and social trust across different NGO–host government relationships"}
sim_final %>% 
  group_by(org_relationship, persona_trust, persona_ideology, persona_experience) %>% 
  summarize(avg_share = mean(share)) %>% 
  ggplot(aes(y = fct_rev(str_wrap_factor(org_relationship, 10)), 
             x = avg_share, color = persona_trust)) +
  geom_pointrange(size = 0.75, fatten = 1.5,
                  aes(xmin = 0, xmax = ..x..), position = position_dodge(width = 0.5)) + 
  scale_x_continuous(labels = percent_format(accuracy = 1), expand = expansion(add = c(0, 0.002))) +
  scale_color_manual(values = clrs_ngo_pairs[[3]], 
                     guide = guide_legend(reverse = TRUE,
                                          override.aes = list(size = 0.25,
                                                              linetype = 0))) +
  labs(x = "Average share of donations", y = NULL, color = NULL) +
  facet_grid(rows = vars(persona_ideology), cols = vars(str_wrap(persona_experience, 100))) +
  theme_ngo() +
  theme(panel.grid.major.y = element_blank(),
        legend.key.width = unit(0.5, "lines"))
```

Figure \@ref(fig:relationship-social) shows how personas respond to different NGO–host government relationships. Public affairs knowledge and experience has little effect on donor preferences, as the average market shares are roughly identical across the two columns of the figure. Persona preferences toward friendly, criticized, and legally suppressed organizations do not change much across different ideological views. The strongest, most overriding factor in determining donor preferences when considering an organization's relationship with its host country is social trust. 

Personas across both levels of social trust are most likely to donate to organizations with friendly relationships with their host governments. However, those with higher trust are less likely to be deterred by poor host-country relationships, while those with low trust tend to not donate to organizations facing criticism or crackdown abroad. A friendly relationship with a host government may be a manifestation of a belief that charities behave properly and exert a positive influence on society. Once an organization faces criticism or crackdown, however, support deteriorates rapidly among personas with low social trust, since such a crackdown could be perceived as a violation of the persona's trust in the charity—that is, the organization must be doing something wrong to deserve criticism. When the risk of a certain action is higher, people need higher levels of confidence or trust to engage in that action [@coleman1990]. When NGOs have goals that are harder to accomplish (as is the case under shrinking civic space), there is greater uncertainty that the donated money will meaningfully contribute to the intended cause. 

Personas with high levels of trust, on the other hand, seem to stick with legally besieged organizations and show targeted groups the benefit of the doubt. The global crackdown on NGOs thus sours NGOs' least trusting individual donors. Overall, our results support research that donors with high levels of trust are more active in giving than those with low levels of trust [@sargeant2008] and that trust especially matters when individuals donate to organizations working under great uncertainty and towards difficult goals [@Wiepking:2010].

## Concerns about inference and generalizability

The conjoint experiment conducted among a nationally representative sample of U.S. respondents with a prior history of charitable giving raises two main concerns for inference. It may be the case that respondent answers—and the corresponding beahavior of simulated personas—could potentially be driven by their priors towards these organizations. In addition, there are also important questions about the generalizability of these results outside the American context. We explore both below.

Given the nature of conjoint experiments, the randomly presented organizations may have influenced respondent choices. All four organizations are well-known, and it is likely that participant responses reflect prior beliefs and news coverage about Amnesty, Oxfam, Greenpeace, and the Red Cross. For instance, in February 2018, news broke out about sexual abuse and misconduct at Oxfam and several other prominent organizations, including Amnesty International and Save the Children. This may have potentially strengthened a narrative that the nonprofit sector suffers from governance failures [@Phillips:2019]. In a competitive NGO environment [@Cooley:2002], maintaining positive perceptions is important as donors who perceive these organizations as untrustworthy may simply take their resources elsewhere. As such, it could be argued that respondents may be less likely to donate to Oxfam or Amnesty because of these scandals. 

However, we remain doubtful that the above news would impact our U.S. respondents' beliefs about these four organizations, as evidence about the effect of these scandals is mixed. Using data from Twitter, @ScurlockDolsakPrakash:2020 find that while Oxfam's reputation suffered for at least six months after the scandal (but has since recovered), other NGOs facing scandals, such as Save the Children, were able to recover more quickly. More broadly, the effect of scandals was mixed—while reputational damage can lead to a decline in donations and grants, organizations can become more resilient and durable after a public relations recovery. 

Additionally, we remain skeptical that this scandal affected U.S. donors, given Oxfam is based in the UK, and "media narratives are often not as unsupportive as often assumed" [@BanksSchulpenBrockington:2020, 703]. Moreover, research shows that when donors feel solidarity or an ideological connection with an NGO – especially because of the cause espoused by them – they are likely to consider the organization to be trustworthy, despite the lack of full transparency and oversight [@KeatingThrandardottir:2017]. Even watchdog organizations such as Charity Navigator and Better Business Bureau Giving Alliance have been found to not significantly impact individuals’ trust in nonprofit organizations. Rather, trust is more likely fostered through local networks and personal scrutiny [@TBPrakash:2017, 643].

We also failed to find any systematic evidence that the Oxfam scandal influenced perceptions of the nonprofit sector in the U.S., likely because the story had far less news coverage in the U.S. Many individual donors have a preference to donate locally and they lack sufficient knowledge to make an informed decision about donating internationally. @KnowlesSullivan:2017, conducting a field experiment in New Zealand, found that only a small number of participants in the study stated that not giving internationally was driven by mistrust. Rather, lack of trust towards overseas NGOs, particularly aid and developmental agencies often results due to donors perceiving INGOs as having low personal relevance [@FaulknerTruongRomaniuk:2015]. Overall, the literature suggests that for INGO fundraising, lack of awareness is a much larger issue than specific scandals or other knowledge of specific organizational characteristics that might reduce donor trust and confidence.

The survey conducted among a random sample of U.S. residents with a prior history of charitable giving also raises questions about generalizability and external validity. Public expectations about the role of government and private funding of the nonprofit sectors differ by country. In the U.S., for instance, NGOs often receive substantial funding through government grants, while those in the UK and France, such as Amnesty International and Médecins Sans Frontières, receive minimal government funding [@Stroup:2012]. This could imply that donors in UK and France may have very different preferences when choosing to donate to government or privately-funded organizations. More comparative research is needed to see if our results would also replicate cross-nationally.

There is also notable country-based variation in general patterns of philanthropy. Though charitable giving is highest in the U.S, such donations are overwhelmingly directed at local and religious causes. This pattern does not hold for many European countries. In the UK, international aid is one of the largest targets of giving (26%). Other countries prioritize international philanthropy even more, such as Germany (74%), Belgium (61%), and Switzerland (43%) [@Milner:2017]. Thus, if crackdown on INGOs abroad can elicit the effects found in this paper among U.S. respondents, the results provide much reason for optimism for pursuing individual philanthropy towards INGOs in Europe as well. However, more comparative research is required to confirm these trends.


# Implications and conclusion

The global crackdown on civil society has adversely impacted the operations of NGOs. It has not just resulted in net losses of income from official sources to INGOs, it has also affected how and where funds are allocated. We argue that individual donors can be useful in sustaining the operations of groups working in challenging environments. Individual donors do not use the same perfomance-based metrics as official aid donors and use heuristics to smiplify their decision-making. The role of trust may be important in understanding donations preferences towards INGOs working in challenging contexts. While trust is a difficult construct to define, the nonprofit literature conceptualizes trust as a key driver of commitment, which is shown to impact both intention to donate as well as actual donation behavior. With repressive environments and NGOs working on more difficult goals, a lack of trust could be further fueled by legal trouble with the NGO host government—a dynamic that individuals may not wish to enter with their dollars. How does trust in NGOs influence individuals' preferences to donate, especially when groups face crackdown?

Using a simulated market for philanthropic donations towards INGOs, based on data from a nationally representative sample of individuals in the United States who regularly donate to charity, we find several important trends. Trust in political institutions and in charitable organizations matters substantially in shaping donor preferences. Personas with high levels of social and charitable trust prefer donating to human rights organizations that have friendly relationships with their host governments, while personas with low trust are drawn more to emergency response and refugee relief organizations with unfriendly relationships with their host governments. Personas with low social trust tend to be wary of negative host-country relationships and strongly prefer to donate to organizations that are friendly with their host governments, while personas with high social trust are more willing to stick with legally besieged organizations. This implies that donors who are the most trusting could remain supportive when nonprofits face government criticism and crackdown. The global crackdown on NGOs may also possibly sour NGOs' least trusting individual donors. Our findings lend support to research showing that individuals who trust NGOs are more likely to continue their monetary support of NGOs in the face of difficulties and under great uncertainty [@dwyer1987; @Wiepking:2010].

Our results have important implications for NGOs as they navigate a shrinking civic space and consider the use of different frames to tailor their fundraising appeals to individual donors. Our simulation finds that at least for some kinds of personas, NGOs may benefit from publicizing when they are targets of government crackdown, including that information in their fundraising appeals. However, the larger challenge may be addressing the concerns of donors with low levels of trust, as they are the least likely to donate organizations facing criticism and crackdown abroad. As @TBPrakash:2017 argue, "trust in the charity sector cannot be assumed: the sector has to demonstrate that the trust it receives is justified." (643).

When considering INGOs working in different issue areas, studies from a range of countries show that levels of giving are lowest for refugee and asylum organizations. Rather, donors giving to international causes tend to support international disaster relief and religious groups [@RobsonHart:2020]. However, some of the former are precisely the organizations facing crackdown. More research is needed to answer how contentious or advocacy organizations (as opposed to service nonprofits) can appeal to broader array of potential foreign donors to increase levels of charitable-giving towards them.

To be clear, research on international giving by individuals, especially in the era of closing civic space, is not meant to find answers that can act as substitutes for strategic policy responses, especially by official aid donors and foundations. However, many NGOs are under immediate threat, and individual-level philanthropy can help support these organizations. Existing initiatives such as the Lifeline Embattled CSO Assistance Funds provide models for channeling such funds. In the face of legal crackdowns abroad, individual donations are an important additional funding source for besieged NGOs.


\newpage

[^smalldonors]: Micro-donations refer to donations between $0.25–$10; small donors are individuals who donate up to $50 [@SaxtonWang:2014].

[^NGOdef]: Broadly speaking, NGOs are any local, national, or international not-for-profit, voluntary organization. They can be domestic (i.e. an NGO that operates entirely in a single state), or international (INGOs). INGOs are composed of members from two or more countries and are organized to advance their members' international goals and provide services to citizens of other states through transactions with states, private actors, and international institutions [@Tarrow:2001, 12].

[^exceptions]: For exceptions, see @Micklewright:2009 on the UK; @Rajan:2009 on Canada; @WiepkingBekkers:2010 on the Netherlands; @CasaleBaumann:2015 on the U.S.

[^highnetworth]: High-net worth households have greater than \$200,000 in annual household income and/or have more than \$1,000,000 in assets, excluding primary residence.

[^phildef]: We use philanthropy to connote voluntary giving to promote the common good. While it can include grants given by foundations and corporations to nonprofit organizations, our focus is on individual giving. Some perspectives see philanthropy as more of a strategic long-term practice addressing the root causes of issues, and charity as a more short-term practice focused on providing immediate relief. But many scholars see philanthropy as essentially synonymous with charitable donations, as the latter can also address longer-term or strategic issues [@Castle:2004]. Despite these subtle differences, we use the terms inter-changeably here both because previous literature does so, and because in the case of giving to NGOs during crackdown, individuals could be giving to nonprofits both for immediate relief as well as long-term strategic purposes. 

[^intlaffairs]: Giving USA defines the sector of international affairs to include any nonprofit organization working in international development, international relief services, disaster relief, international human rights, international peace and security, foreign policy research and analysis, and international exchange programs.

[^sample-size]: A sample size of at least 500 respondents is typical when using conjoint data in a hierarchical Bayesian model. We doubled this amount because we are interested in analyzing subpopulations of respondents, which requires a larger sample.

[^appendix-survey]: See the appendix for the full text of the survey.

[^model-details]: Specifically, 

    $$
    \begin{aligned}
    \beta &\sim \operatorname{Multivariate} \mathcal{N}(Z \Gamma, \xi) \\
    y &\sim \operatorname{Multinomial\ logit}(X \beta, \varepsilon)
    \end{aligned}
    $$

    where $y$ = which alternative the respondent chooses to donate to, $X$ = design matrix of attribute levels (i.e. organizations, issue areas, organizational practices, funding sources, and government relations), $\beta$ = latent individual preferences for the attribute levels, $Z$ = matrix of individual-level covariates (i.e. demographics, political knowledge, attitudes towards charity, etc.), $\Gamma$ = matrix of coefficients mapping individual-level covariates onto the latent individual-level preferences, and $\varepsilon$ and $\xi$ = errors.
