---
title: "Who Cares About Crackdowns? Exploring the Role of Trust in Individual Philanthropy"
short-title: "Who Cares About Crackdowns?"
author:
- name: Suparna Chaudhry
  affiliation: Lewis & Clark College
  url: http://www.suparnachaudhry.com/
  email: schaudhry@lclark.edu
- name: Marc Dotson
  affiliation: Brigham Young University
  url: https://marriottschool.byu.edu/directory/details?id=50683
  email: marc_dotson@byu.edu
- name: Andrew Heiss
  affiliation: Georgia State University
  url: https://www.andrewheiss.com/
  email: aheiss@gsu.edu
date: "June 8, 2020"
published: Working paper.
git-repo: https://example.com
abstract: >-
  Over the past two decades, international nonprofit organizations (INGOs) have faced increasing legal scrutiny abroad, which has negatively impacted public funding flows. Consequently, INGOs may feel the need to fundraise from individuals. Which demographic and experiential factors influence the decision of individuals to donate to legally besieged NGOs? We investigate this question using a market simulation based on a nationally representative conjoint survey experiment conducted on individuals with a history of charitable giving. We find that trust in political institutions and charities is enormously influential in shaping donor preferences. Donor profiles with high levels of social trust are likely to donate to human rights INGOs with friendly relationships with host governments, but that support evaporates if INGOs face criticism or crackdown. In contrast, donor profiles with lower levels of social trust prefer to donate to organizations that do face criticism or crackdown abroad. Our findings have practical implications for INGOs that hope to raise funds from the general public amid closing civic space. 
reference-section-title: References
toc: false
mainfont: Lora
sansfont: IBM Plex Sans
---

<!-- manuscript.md is generated from manuscript.Rmd. Only edit the .Rmd file, *not* the .md file. -->

In Human Rights Watch's 2016 annual report, Executive Director Kenneth Roth argued that civil society was under more aggressive attack than at any time in recent memory. Over the past few decades, governments across the globe have been cracking down on civil society organizations (CSOs) and non-governmental organizations (NGOs)—a phenomenon known as "closing civic space" in the policy arena [@CarothersBrechenmacher:2014; @CIVICUS:2017].[^NGOdef] As part of this phenomenon, CSOs, particularly NGOs, face increasing obstacles to their work, including legal restrictions and threats to their freedom of association, expression and speech. Legal restrictions-a core part of these efforts-create barriers to entry, funding, and advocacy for NGOs in an effort to control, obstruct or repress these organizations-what we refer to as legal crackdowns in this paper. 

As a result of these legal crackdowns, official aid flows to NGOs have decreased considerably in repressive countries [@Brechenmacher:2017; @ChaudhryHeiss:2018; @DupuyPrakash:2018]. While this has negatively impacted funding flows from public organizations like USAID, DFID, and other OECD funding agencies, philanthropy from private donors, in particular, foundations, appears to not be as adversely affected [@McGill:2018]. Given this trend of closing civic space, NGOs need access to multiple sources of funding and individual-level giving can serve as one such source of funds. What individual donor characteristics and attributes are most salient for INGOs interested in maximizing philanthropy towards their operations abroad?

Existing scholarship on the determinants of individual-level philanthropy to domestic causes is robust across a variety of contexts. But the dynamics of *international* giving to NGOs are under-examined in the literature on individual donor attributes and preferences.[^exceptions] Research on the determinants of philanthropy towards international causes—particularly for organizations working in challenging environments—is still in its infancy due to the absence of expansive and high-quality data [@greenhill2013]. Understanding why individuals donate to international causes is essential as it can be harder to motivate as the number of recipients are large and far removed from the donor [@CasaleBaumann:2015, 100].

Individual giving is also important to understand because the amount of money donated by private individuals to CSOs working abroad has steadily increased over time. Total aggregate philanthropy in the United States has grown substantially over time, reaching more than \$400 billion in 2018 (@fig:giving-aggregate). In 2018, total private giving to CSOs working specifically in international affairs reached \$22.08 billion. Moreover, organizations working in the area of international affairs are one of the only two types of recipient organizations that have continued experiencing substantial growth in giving [@givingusa:2019].[^intlaffairs] Most of this is driven by individuals making small-scale or micro-donations.[^small donors] In 2017, individuals gave \$286.7 billion, or approximately 70% of total giving to organizations working in international affairs. This further increased to \$427.71 billion in 2018, 85.7% of which was through individuals [@givingusa:2019]. However, we lack a systematic understanding of which individual donors may be further motivated or deterred to donate under an increasingly restricted civic space abroad. What is the optimal mix of individual-level donor attributes for international NGOs working in restricted political environments interested in maximizing donations? What kinds of individual donors are more or less likely to consider an NGO's host country political environment when making the decision to donate? 

![Aggregate philanthopic giving in the United States, 1977–2017. Source: Giving USA.](output/figures/giving_aggregate.pdf){#fig:giving-aggregate}

We explore these questions using a simulated market for philanthropic donations, based on data from a nationally representative sample of individuals in the United States who regularly donate to charity. We adapt methods used in marketing research to generate dozens of distinct simulated personas that represent typical donor profiles, and we vary these personas along different dimensions of demographic, political, and social attributes. We find that specific individual characteristics help shape donor preferences for donating to international NGOs, and that an INGO's target issue area and its relationship with host government interact strongly with these characteristics. Personas with high levels of trust in political and charitable institutions and who are actively engaged in charity and volunteerism prefer donating to human rights organizations that have friendly relationships with their host governments. In contrast, personas with low levels of social trust eschew both human rights organizations *and* any type of NGO with friendly relationships with host governments, preferring instead to donate to NGOs facing more restrictions and working on less contentious issues like emergency response and refugee relief. This may indicate that donors with high trust and support for NGOs could be fair-weather friends that drop their support when their preferred organizations meet resistance abroad.

This study contributes to both to scholarship and practice in public policy, nonprofits and charitable-giving. Individual-level philanthropic giving abroad can help bridge socioeconomic disparities worldwide, especially given the stark inequality in global distribution of income. Knowledge of demographic traits and its impact on funding can help nonprofits craft more effective and targeted marketing strategies [@Rajan:2009, 414]. Most NGOs working internationally have previously relied on government and foundation funding. But in the era of closing civic space, these groups may need to readjust their fundraising strategies to cater to individuals. Framing is important in this regard—how people process information affects their willingness to donate.[^framing] However, we lack systematic studies of which frames may be effective when making appeals to individual donors considering international philanthropy. Since our study looks at multiple organizational-level factors, it can help these CSOs frame their appeals.

More broadly, while research in public policy and nonprofit studies has focused on giving to NGOs, philanthropy is not seen as relevant by international relations as a discipline [@youde2019]. However, philanthropy—particularly donations received by international nongovernmental organizations (INGOs)—can help us understand changes in the dynamics of global governance. Private individuals exercise power and influence in unique ways that differ from other non-state actors. As @youde2019 [p. 44] points out, it changes our conceptions of individuals being passive recipients to active entities that are not merely objects of governance, thus shedding light on the changing logic of global governance.

Below, we first examine the state of the existing literature on the key drivers of individual charitable giving, as understanding what motivates private donations is crucial to more effective NGO fundraising. We then describe our survey experiment and simulation methods and explore the results. We conclude with a discussion of the implications our findings have on general INGO funding in the age of constrained civil society space.


# Sustaining civil society organizations abroad

Foreign aid to civil society—and NGOs in particular—has steadily increased over time. Many government and foundation donors perceive that NGOs have a superior ability to achieve intended outcomes. Therefore, in countries with weak institutions and poor governance, where direct aid transfers to governments raise issues of misuse and bureaucratic inefficiency, donors specifically seek out NGOs for projects [@Dietrich2013]. These groups provide services to large parts of the population in the Global South, especially where governments are unable to do so on a consistent basis. 

## Sources of funding

Thousands of INGOs in the Global South receive funds from a variety of public and private sources each year. Public funds (more commonly known as traditional donor aid) channel money from official aid agencies through a variety of bilateral and multilateral institutions. Private sources include foundations, private corporations, and individuals. While INGOs have not been passive in responding to the crackdown on civil society, they face increasing obstacles especially from a funding perspective. Official aid donors have reduced the amount of aid channeled to countries with legal restrictions @DupuyPrakash:2018]. In addition, official donors have also redirected funds within restrictive countries, decreasing funds for politically sensitive causes in favor of tamer issues that are more compatible with the governments of other countries [@ChaudhryHeiss:2018]. However, it is notable that private funding towards international nonprofits has not been as adversely affected. Many foundations have continued channeling funds to countries with hostile legal environments [@FoundationCenter2018].

The advent of funding restrictions on NGOs globally and a subsequent decline in government and foundation funding has made the role and value of individual philanthropy in sustaining NGOs even clearer.[^phildef] NGOs need access to a diverse array of funds and individual donations to international affairs organizations have steadily increased over the past two decades. Between 2014-2016, giving to organizations working in international affairs increased by 14.1% [@givingusa2017report]. Moreover, at least in the U.S. context, donations from individuals far outstripped foundation and corporate giving. In 2018, individual donations amounted to \$292.09 billion, compared to \$75.86 billion from foundations and \$20.05 billion from corporations [Source from NVSQ]. Given that a majority of research is restricted to elite high net-worth donors [@USTrust2014], these figures show that it is imperative to understand motivations and attributes of small-scale donors as well. In this paper, we therefore look at which kinds of individual donors—beyond high-net worth donors—are most likely to give abroad.[^highnetworth]

## Key drivers of individual-level philanthropy

A vast literature on philanthropy looks at organizational-level factors that motivate individuals to donate [@BilodeauSlivinski:1997; @TBPrakash:2017]. However, donation decisions are not determined solely by an organization's characteristics. Donors have personality traits, preferences, and experiences that make them more or less likely to engage in philanthropy. A robust literature in nonprofit studies has looked at individual donor behavior and characteristics for giving to local—rather than international—causes. This literature focuses on the importance of five main attributes: (1) social capital; (2) the role of higher education; (3) levels of religiosity; (4) wealth; and (5) national group attachment.

One major area of inquiry in philanthropic studies examines the role of social capital, particularly associational capital, or "the networks of formal and informal relationships to which people are associated." [@havens2007, 240]. Research finds that those who participate in a variety of associations are more likely to make larger donations [@hossainlamb2017]. Further, those who volunteer their time (compared to those who did not) were both more likely to make charitable donations and make bigger donations. Volunteering exposes individuals to the need for donations, making them feel more compelled to donate [@hossainlamb2012]. 

Higher education can also play an important role in charitable-giving decisions as it fosters prosocial motivations. It also brings people into particular social networks that entail a higher level of solicitations, which is vital in giving to international causes [@BekkersWiepking:2011]. @Micklewright:2009 show that higher education matters more for donating to international relief organizations than donating to domestic organizations. 

Higher education also helps determine the kinds of causes that people donate to globally. A number of NGOs facing crackdowns abroad work on issues that may be seen as "contentious," including human rights, elections, corruption, advocacy, media freedom—issues that may have a harder time attracting funds due to their non-essential nature when compared to causes like health, education, sanitation, and social services. However, additional education has been shown to increase abstract thinking [@wiepkingmaas, 1978], which may be necessary to donate to causes that do not have quantifiable goals, which is particularly important for organizations with more "contentious" causes as part of their missions [@Bush:2015]. 

Levels of religiosity are also a significant predictor of the likelihood and level of giving. People who regularly attention religious services are more likely to have prosocial values, face a stronger norm of charitable giving, and may feel pressure to conform that group and community standard [@wiepking2010]. Regular attendance may also strengthen individuals' beliefs in organizations' missions. 

Finally, recent research looks at how a sense of ethnonationalism and national group attachment affects the decision to donate to domestic and international charities. Looking at UK respondents, @hartrobson2019 find that whereas nationalists have a preference for domestic charities, they are not necessarily averse to donating to international causes, whereas internationalists have a strong preference for donating to causes and groups working abroad. More specifically, immigrant households are more likely to contribute to international aid organizations [@OktenOsili:2007], and that immigrants can find value in donating to both home and host cultures, especially if their family members still reside in the home country [@burgoyne2005]. 

Overall, individuals with higher levels of income, education, and greater religious proclivities are not only more likely to give internationally, but also give higher amounts [@Rajan:2009; @CasaleBaumann:2015]. While we expect these determinants to be reflected in our survey experiment, we do not lay out specific hypotheses. Because there is so little research about international donations during legal crackdown by host governments, our analysis and results are exploratory and descriptive. Our innovative experimental approach, however, provides rich quantitative details of individual and organizational characteristics simultaneously and reveals unique insights into donor motivations that both confirm and expand previous research.


# Data and methods

## Conjoint survey experiment

We use a conjoint survey experiment to simultaneously study the effects of organizational characteristics and individual attributes on individuals' preferences to donate to a nonprofit organization. Conjoint analysis methods are well suited for the questions we explore, as they allow us to elicit donor preferences without forcing respondents to explicitly state their exact preferences. Conjoint analysis is commonly used in marketing research to reveal consumer preferences for different mixes of products. It has also been increasingly used in political science and public policy to assess citizen perceptions of political communication [@Helfer:2016; @Mummolo:2016; @KnudsenJohannesson:2019]. Because we are interested in exploring the complexity behind the decision to donate, we apply conjoint methods to the study of nonprofits and philanthropy.

After collecting baseline information on respondent demographics and attitudes toward charity and voluntarism, we presented respondents with four randomly selected international nonprofit organizations with four randomly assigned features.^[See the appendix for a copy of the survey.] We included organizations that are associated with a range of contentious and noncontentious issues: Amnesty International, Greenpeace, Oxfam, and the International Committee of the Red Cross. These organizations are commonly used in surveys on TODO TODO, in addition to other organizations such as MSF, World Vision (LOOK UP SAVED PAPERS - I THINK NVSQ). We then displayed various combinations of features, including one of four issue areas (emergency response, environmental advocacy, human rights advocacy, and refugee relief), one of two organizational practices (financial transparency and third-party accountability audits), one of three funding sources (many small private donations, a handful of wealthy private donors, and government grants), and one of three relationships with host governments (friendly, criticized by the government, and under government crackdown). We presented twelve sets of four hypothetical combinations of organizational features, and because our partial fractional factorial design resulted in 288 (4 × 4 × 2 × 3 × 3) possible combinations of organizations and features, no single respondent was offered every combination.

In June 2019, we recruited survey respondents through Centiment, a commercial provider of high quality non-probability opt-in survey panels. Centiment ensures panel quality by actively recruiting representative samples of the US population and rewards participants with monetary incentives. To participate in the study, respondents were first screened based on their charitable activities. Those who indicated that they gave to charity once every few years or never were disqualified. After screening, we collected a sample of 1,016 respondents, which is sufficient for model estimation.[^sample-size]

\newpage

```{r child = "output/tables/tbl-sample-details.md"}
```

@tbl:sample-details provides a summary of the individual characteristics of survey respondents. Because we limited the sample to people who regularly make charitable contributions, most respondents have a favorable opinion of nonprofits and charities, with more than half reporting a 5 or greater level of trust in charities (on a 7-point scale). Moreover, nearly 90% indicated that it is important to trust charities. Additionally, the majority of respondents volunteer at least once a year, donate to charity at least once a year, and give between \$1–\$999 annually. Most rarely attend church, but feel that religion is important. Respondents are also civically engaged (three quarters voted in the last election) and follow national news closely (70% follow the news daily). Respondents are fairly balanced ideologically, with the majority reporting a 3, 4, or 5 on a 7-point scale of liberal–conservative political ideology.

We analyze the effect of different combinations of organizational features and individual respondent attributes with a two-stage hierarchical Bayesian multinomial logit model.[^model-details] Simply put, the first level of the model predicts individual donor preferences for various combinations of features (e.g. the combination of human rights issues, financial transparency, government funding, and government crackdown), while the second level of the model uses these fitted preferences to predict the ultimate choice of donation.

## Simulated philanthropy market

In this paper, we look at which kinds of donors are more or less likely to consider an NGO's issue area, funding sources, and host country politics when deciding to donate. Disentangling the effect of every individual level-characteristics measured in the survey on every combination of organizational characteristics presented in the conjoint experiment is unwieldy and near impossible. Given the organizational attributes presented in our experiment, and the possible individual characteristics that respondents could select, there are billions of possible combinations and iterations of organizational and individual attributes—many of which rarely occur in the actual population of potential donors. To more fully explore the relationships between different organizational- and individual-level attributes, we turn to simulation methods that allow us to examine the impact of multiple covariates at simultaneously

We use market segmentation methods to construct simulated personas that interact with simulated organizations in a market for charitable contributions. To simplify the many possible combinations of individual characteristics, we generate 32 distinct personas with attributes that vary along different dimensions of demographics, politics and public affairs experience, and social attitudes (see @tbl:persona-attributes). We also generate 24 simulated organizations that vary by issue area, relationship with host government, and source of funding (see @tbl:organization-attributes). 

```{r child = "output/tables/tbl-persona-attributes.md"}
```

```{r child = "output/tables/tbl-organization-attributes.md"}
```

We then combine these different scenarios of donor and organization attributes with the coefficients from a multinomial multilevel regression model to calculate the probability that each simulated persona has of giving to specific combinations of organizational attributes. Any individual characteristics that we do not explicitly vary in  @tbl:persona-attributes are held at their means or modal values from the survey sample. @tbl:sim-output provides an example of the simulation output, showing how two different personas are predicted to allocate donations to organizations with various characteristics. The proportion of donations from each persona sums to 100%, as every persona is guaranteed to make a donation. 

```{r child = "output/tables/tbl-sim-output.md"}
```

Relying on simulation methods offers useful analytic benefits. Rather than interpret individual marginal effects from a regression model, we can explore multiple moving parts from the same model simultaneously to understand how individual characteristics interact with each other. This allows us to more fully explore our key research questions and see which kinds of donors are more or less likely to consider specific organizational attributes when deciding to donate, and which donor characteristics are associated with the propensity to donate. Importantly, each persona is an artificial construct. While some survey respondents might match an exact persona profile, most do not. This is acceptable in simulations, and each combination of persona characteristics is plausible in the real world. Our analysis uses the characteristics of our sample to model the effects of various salient attributes while holding others constant, and thus allows us to explore multiple attributes at once. 

Also importantly, our findings are not causal. Rather, they are the descriptive results of simulated data from a nationally representative sample of people who regularly donate to charity. The insights from this simulation, however, still reveal interesting patterns of donor preferences when faced with organizations focusing on a variety of issues, funding sources, and legal restrictions abroad.


# Results

Our market simulation reveals which combination of organizational features each persona is most likely to express a preference for donation to. We examine the results of the simulation in three stages, looking at how personas with different (1) demographic characteristics, (2) public affairs knowledge and experience, and (3) levels of trust in government, charitable work, and associational life change their donation preferences across different organizational issue areas, relationships with host governments, and funding sources.

## Demographics

We begin by looking at differences in donation preferences across a range of basic demographic characteristics. The three panels of @fig:demographics shows the average predicted market share for different organizational issue areas, relationships with governments, and funding sources. Though we varied gender across the different personas, gender has no significant effect on any of the simulated outcomes—males and females follow nearly identical patterns across all organizational and individual characteristics. As such, in @fig:demographics, we only vary income. Personas with high income earn \$100,000 a year, while those with less income earn \$50,000 annually.

![Average predicted donation market shares across all personas, segmented by individual persona income and organizational issue area, relationship with host government, and funding sources](output/figures/income-all.pdf){#fig:demographics}

Wealthier personas respond better to human rights organizations over refugee assistance and environmental advocacy organizations, while personas that earn less money tend to prefer donating to refugee and environmental organizations. As we explore below, this is likely because income correlates with interest in public affairs and political activity—individuals who are wealthier arguably have more time and resources to follow the news and be continuously engaged in their communities and in other activist causes. 

There are also sizable income-based differences across different NGO host-country relationships. Wealthier donors are more likely to prefer donating to organizations that are criticized by their host governments. This falls in line with their preference for human rights organizations which have more contentious missions and are more likely to receive government criticism. This support, however, appears to evaporate as civil society crackdown intensifies. High income personas are the least likely demographic to prefer donating to NGOs under crackdown. Their lower income counterparts, on the other hand, are generally unfazed by NGOs' relationships with government, maintaining the same average market share across all three types of host country relationships. Finally, as seen in the right panel of @fig:demographics, there are no sizable differences in donation preferences by income across different forms of NGO funding. Income thus appears to influence preferences particularly for NGOs facing criticism and crackdown abroad—wealthier donors are perhaps fair-weather friends that are more likely to pull their support when an organization's work becomes more difficult.

*(Note: We also plan on examining religiosity and education in the very near future)*

## Public affairs knowledge, experience, and social trust

We next look at how differences in public affairs knowledge, experience, and political ideology influence a persona's propensity to donate to an organization. We vary personas along two dimensions: (1) political ideology, with semi liberal and semi conservative political leanings (2 and 5 on a 7-point scale), and (2) knowledge and experience with public affairs and the news, with personas that either follow national and international news and that have travelled abroad or personas that do not follow the news or travel. 

We also combine these dimensions with an additional set of donor characteristics that vary a persona's level of social trust. Personas with high levels of trust have high trust in political institutions and charities, thinks people should be more charitable, frequently volunteers, donates to charity once a month, has a history of political activism, and is a member of an association. Personas with low levels of trust have the opposite characteristics: they do not trust political institutions or charities, rarely donate or volunteer, and are not members or other associations. 

![Average predicted donation market shares across all personas, segmented by persona public affairs knowledge, political ideology, and social trust across different NGO issue areas](output/figures/issue-social.pdf){#fig:issue-social}

@fig:issue-social, @fig:relationship-social, and @fig:funding-social show the results of the market simulation across these three dimensions of persona characteristics, varying public affairs knowledge, political ideology, and social trust. Personas respond differently to organizational issue areas. In @fig:issue-social, those who follow the news and travel abroad are more likely to prefer donating to human rights organizations than their less knowledgeable and less traveled counterparts. Personas who do not follow the news or travel show a greater preference for emergency response and refugee relief organization, similar to the income effects seen previously. There are no strong ideological effects across issue areas. Both conservative and liberal personas tend to follow similar patterns of preference across NGO issue. 

Social trust plays a substantial role in determining simulated market share, however. Those who trust social and charitable institutions are very likely to prefer donating to human rights organizations, regardless of their public affairs knowledge or political leanings. In all circumstances, those with high levels of trust have between a 7–10% chance of donating to a human rights organization, on average. Personas with low levels of trust, on the other hand, strongly avoid human rights organizations and prefer instead to donate to refugee and emergency response NGOs. The difference is substantial—a liberal persona with high levels of public affairs knowledge and experience will give to a human rights organization 10% of the time, on average if they have high levels of trust, but less than 3% of the time if they are less trusting. Trust in charities and institutions is a key factor in deciding to donate to more contentious human rights organizations.

![Average predicted donation market shares across all personas, segmented by persona public affairs knowledge, political ideology, and social trust across different NGO–host government relationships](output/figures/relationship-social.pdf){#fig:relationship-social}

@fig:relationship-social shows how personas respond to different NGO–host government relationships. Public affairs knowledge and experience has little effect on donor preferences, as the average market shares are roughly identical across the two columns of the figure. Similarly, there are no ideological differences across the rows of the figure. Persona preferences toward friendly, criticized, and legally suppressed organizations do not change much across different ideological views. The strongest, most overriding factor in determining donor preferences when considering an organization's relationship with its host country is social trust. Personas with high levels of social trust are overwhelmingly most likely to donate to organizations with friendly relationships with their host governments, while those with low levels of trust avoid friendly NGOs almost entirely. 

This stark difference in attitudes towards friendly NGOs likely reflects the underlying levels of trust in the model. For those who trust charities to do the right thing and have a positive influence on society, a friendly relationship with a host government is a manifestation of that belief and a sign that the organization is living up to its idea. Once the organization faces criticism or crackdown, however, support deteriorates rapidly, since such a crackdown could be perceived as a violation of the persona's trust in the charity—that is, the organization must be doing something wrong to deserve criticism. Personas with low levels of trust, on the other hand, eschew friendly organizations entirely and are far more willing to donate to organizations facing criticism and crackdown, likely because they do not assume that charities must always be above reproach. The global crackdown on NGOs could thus counterintuitively sour NGOs' most trusting donors.

![Average predicted donation market shares across all personas, segmented by persona public affairs knowledge, political ideology, and social trust across different forms of NGO funding](output/figures/funding-social.pdf){#fig:funding-social}

Finally, @fig:funding-social shows the average share of persona market shares across different forms of NGO funding. There are only insignificant differences in donor preferences across public affairs knowledge and experience, political ideology, and social trust. On average, it appears that the source of NGO funding does not play an important role in a donor's preference for the organization. Other organizational factors—in particular issue area and host government relationship—have a much larger influence in shaping donor preferences.

## Concerns about inference and generalizability


While the survey seeks to uncover donor profiles of those who are influenced by legal crackdowns on NGOs in their decision to donate, it may be possible that results are based on respondents' prior beliefs about Amnesty, Oxfam, Greenpeace, and Red Cross. This may particularly be the case after an Oxfam scandal in Feburary 2018—scandals that may potentially strength the narrative that the nonprofit sector suffers from governance failures [@SurlockEtal2020]. Analyzing a smaple of Twitter data to studythe daily gap between the positive and negative sentiments expressed towards INGOs embroiled in scandals, @SurlockEtal2020 find that while sentiments towards Oxfam did not completely recover even 6 months after the scandal, Save the Children's reputation recovered fairly quickly. Moreover, the recovery of their reputation was aided specifically by celebrity endorsements as well as be focusing public attention on the organizations' achievements. While reputational damage can lead to decline in donations and grants, their research lays provides important foundational evidence demonstrating both, "short-term fragility and longer-term durability of organizations' reputations" [SurlockEtal2020, p. 104].

However, given that the above analysis is not representative of the general population, we also sought out other surveys that could provide an insight into donors' prior beliefs and their possible impact on donation preferences. Contrary to media sentiment, we found that trust and confidence in NGOs specifically in the UK did not waver [@citemehere]. We also failed to find evidence that these impacted public opinion among likely individual donors in the US - who were the main respondents in our survey [@CITE]. Rather, any lack of trust towards overseas NGOs, particularly aid and developmental agencies often results due to donors perceiving INGOs as having low personal relevance [@FaulknerTruongRomanuik2015]. Many individual donors have a preference to donate locally and they lack sufficient knowledge to make an informed decision about donating internationally. Overall, the literature suggests that lack of awareness is a much bigger issue for INGO fundraising, rather than specific scandals that would reduce the average individual donors' trust and confidence in these NGOs. 

The study was also conducted on a random sample of U.S. respondents with a prior history of charitable giving which raises questions about external validity. Public expectations about the role of public and private sectors differ across the U.S. and European countries. In the U.S., CSOs are typically seen as a counterweight to the government, whereas in some European countries like Germany, states are seen as supporting CSOs. Further, when we consider large INGOs such as Amnesty International and MSF - both receive very little support from their home governments (Stroup 2012). This may imply that donors in UK and France may have very different preferences when choosing to donate to government or privately-funded organizations. As such, more comparative research is needed to see if these results would also replicate across a variety of European countries

There is also notable variation in terms of the targets of donations across various contexts. Though charitable-giving amounts are highest in the U.S, such donations are overwhelmingly towards local and religious causes. However, this pattern does not hold for many European countries. In the UK, international aid is often the largest component of giving (26%) while in Germany, Belgium, and Switzerland, international and humanitarian aid is considered a priority with 74%, 61% and 43% of donations are directed towards these causes in these countries, respectively. Thus, if crackdown on nonprofits abroad can elicit the results demonstrated in this paper among US respondents, then they provide reasons for optimism for pursuing individual philanthropy in Europe as well. However, more comparative research is required in order to confirm these trends.


# Implications and conclusion

The global crackdown on civil society and the consequent declining official aid and official development assistance has made the role of philanthropy in sustaining CSOs abroad clear. INGOs need access to diverse and multiple sources of fundings and the value of individual-level philanthropy is still relatively under-explored. While existing research has examined the impact of organizational-level factors on individual giving, in this paper, we specifically look at demographic traits of those who are likely to donate to INGOs facing hostile environments abroad. Based on our simulated market of a nationally representative sample of individuals from the U.S., we find several important trends. On average, human rights organizations tend to attract personas who are wealthier, follow national and international news, and travel abroad. Donors who are less wealthy, follow the news less often, and do not travel are more attracted to organizations working on less contentious like emergency response and refugee relief. 

More importantly, we found that trust in political institutions and in charitable organizations matters substantially in shaping donor preferences. Personas with high levels of social and charitable trust prefer donating to human rights organizations that have friendly relationships with their host governments, while personas with low trust are drawn more to NGOs emergency response and refugee relief organizations with unfriendly relationships with their host governments. Personas with low social trust rarely prefer to donate to any type of organization with friendly host-country relationships, while personas with high social trust tend to avoid organizations with unfriendly relationships. This could imply that donors that are the most trusting and involved might be fair-weather friends that drop their support when charities face government criticism and crackdown.

This survey was conducted on a random sample of US respondents. More comparative research is needed to understand the extent to which these results can be replicated across other countries. Future research should also determine the extent to which these preferences translate into behavior.

Surveys across a range of countries show that levels of giving are lowest for refugee and asylum organizations, as well as for overseas and development organizations. Rather, donors giving to international causes tend to support international disaster relief and religious groups [@RobsonHart2020]. However, the former are precisely the organizations facing crackdown. More research is needed to answer how can these organizations appeal to broader array of potential foreign donors to increase levels of charitable-giving towards them? To be clear, research on international giving by individuals, especially in the era of closing civic space, is not meant to find answers that can be substitutes for strategic policy responses, especially by official aid donors and foundations. However, many NGOs are under immediate threat, and individual-level philanthropy can help support these organizations. Existing initiatives such as the Lifeline Embattled CSO Assistance Funds provide models for channeling such funds. To be sustainable in the era of closing civic space, CSOs need access to diverse and multiple sources of funding and this paper highlights one such source. 

\newpage

# Appendix: Survey experiment

```{r child = "experiment.md"}
```

\newpage

[^NGOdef]: Broadly speaking, NGOs are any not-for-profit, voluntary group organizations at the local, national, or international level. They can either be domestic i.e. an NGO that operates entirely in a single state, or they can be international, more commonly known as INGOs. The latter are composed of members from two or more countries and are organized to advance their members' international goals and provide services to citizens of other states through routine transactions with states, private actors, and international institutions [@Tarrow:2001, 12].

[^exceptions]: For exceptions, see @Micklewright:2009 on the UK, @Rajan:2009 on Canada, @wiepking2010 on Netherlands, @CasaleBaumann:2015 on the USA. 

[^highnetworth]: High-net worth households are defined as having greater than \$200,000 in annual household income and/or with more than \$1,000,000 in assets, excluding primary residence.

[^phildef]: We use philanthropy to connote voluntary giving by an individual or group to promote the common good. While it can include grants given by foundations and corporations to nonprofit organizations, our focus here is on individual giving. 

[^framing]: Framing is the process through which actors present information to influence perceptions or behavior. Previous research has shown that which frames are employed can have a significant effect on individuals' attitudes and behaviors [@BenfordSnow:2000; @ChongDruckman:2007].

[^intlaffairs]: Giving USA defines the sector of international affairs to include any nonprofit organization working in international development, international relief services, disaster relief, international human rights, international peace and security, foreign policy research and analysis, and international exchange programs. 

[^sample-size]:	A sample size of at least 500 respondents is typical when using conjoint data in a hierarchical Bayesian model. We doubled this amount because we are interested in analyzing subpopulations of respondents, which requires a larger sample.

[^model-details]:	Specifically, 

    $$
    \begin{aligned}
    \beta &\sim \operatorname{Multivariate} \mathcal{N}(Z \Gamma, \xi) \\
    y &\sim \operatorname{Multinomial\ logit}(X \beta, \varepsilon)
    \end{aligned}
    $$

    where $y$ = which alternative the respondent chooses to donate to, $X$ = design matrix of attribute levels (i.e. organizations, issue areas, organizational practices, funding sources, and government relations), $\beta$ = latent individual preferences for the attribute levels, $Z$ = matrix of individual-level covariates (i.e. demographics, political knowledge, attitudes towards charity, etc.), $\Gamma$ = matrix of coefficients mapping individual-level covariates onto the latent individual-level preferences, and $\varepsilon$ and $\xi$ = errors. 
